---
title: "Simple document"
author: "Chelsey Oppan"
date: 2025-09-11
output: html_document
---

I'm an R Markdown document! 

R code chunk shortcut - option + command + I

```{r echo = TRUE, message = FALSE}
library(tidyverse)
```
shows code but hides message you write

```{r eval = FALSE}
library(tidyverse)
```
shows the code but does not evaluate it

```{r}
library(tidyverse)
```

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```
click "knit" to create R markdown document

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`. (in line R code - allows you to do R code interspersed within in text)

# Section 3

This is where I'm going to talk about code chunks.

```{r}
mean(samp)
sd(samp)
```

let's also make a data frame.
```{r}
example_df = 
  tibble(
    vec_numeric = 1:4,
    vec_char = c("my", "name", "is", "chels"),
    vec_factor = factor(c("male", "male", "female", "female"))
  )
```

I'll create a new data frame

```{r}
new_df = 
  tibble(
    x = rnorm(100),
    y = 1+2 * x + rnorm(100)
  )
```

Let's make a plot and see how cool that is!
```{r}
plot_df = 
  tibble(
     x = rnorm(100, mean = 1, sd = .5),
    y = 1 + 2 * x + rnorm(100)
  )

ggplot(plot_df, aes(x=x, y=y,)) + geom_point()
```

# Section 4 

This chunk is a learning assessment.

```{r}
learning_df = 
  tibble(
    x = rnorm(500, mean =1),
    vec_log = x > 0,
    vec_numeric = abs(x)
  )

ggplot(learning_df, aes(x=vec_numeric)) + geom_histogram()

```




